### 2023-03-02 figures/first_layer.tex
add figures for first layer impulse response

\begin{figure*}[htb]
	% blog post on how to use pgf from matplotlib to latex
	% https://timodenk.com/blog/exporting-matplotlib-plots-to-latex/
	% \usepackage{layouts} and \printinunitsof{cm}\prntlen{\textwidth} show that textwidth is 17cm here.
	\centering
	\input{figures/first_layer.pgf}
	\caption{First layer impulse responses.}
	\label{fig:first_layer}
\end{figure*}