### 2023-03-08 figures/first_layer.tex
update figure and corresponding text

\begin{figure*}[htb]
	% blog post on how to use pgf from matplotlib to latex
	% https://timodenk.com/blog/exporting-matplotlib-plots-to-latex/
	% \usepackage{layouts} and \printinunitsof{cm}\prntlen{\textwidth} show that textwidth is 17cm here.
	\centering
	\input{figures/first_layer.pgf}
	\vspace{-0.25cm}
	\caption{Frequency responses for (parts of) the different \fes. (a)-(d) show the frequency response of the filters in the first layer that operates on the waveform. (e)-(f) show the frequency response of the complete learned \fe.}
	\label{fig:first_layer}
\end{figure*}